```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Binomial generalised linear model analysis of risky choices  
## Section 1  

### Introduction  
In this study, we examined the effects of framing on risky choice using a binomial generalised linear model with its canonical link, the logit on a within-subjects design. The main research aim was to determine if a framing effect exists in which more risky choices are made under loss frames as compared to gain frames and if the effects differ across the three samples. A minor research aim was determine if participants were generally risk averse. The independent variables are sample with 3 levels (Connecticut College, Malaysia and Prague) and framing with 2 levels (gain and loss). The dependent variable is risky choice - the binary response of participants for safe and risky options. Follow-up analyses controlled for type I errors using the Bonferroni-Holm method comparing the gain and loss frame cells for each of the samples.  
  
### Summary  
  
A main effect of sample ***F*(2, 267) = 3.64, p < 0.05** was found. The main effect of framing was also significant, ***F*(1, 267) = 16,64, p < 0.0001** but the interaction of sample with framing was not significant ***F*(2, 267) = 2.81, p > 0.05**. 

```{r include = FALSE}
library("tidyverse")
library("afex")
library("emmeans")
theme_set(theme_bw(base_size = 15) + 
            theme(legend.position="bottom", 
                  panel.grid.major.x = element_blank(),
                  panel.grid.minor.x = element_blank()))
options(digits = 3)
set_sum_contrasts()
frame <- read_csv("framing_replication.csv")
frame <- frame %>% 
  mutate(sample = factor(sample, 
                         levels = c("ConnecticutCollege", 
                                    "Malaysia", 
                                    "Prague")),
         framing = factor(framing,
                          levels = c("gain", "loss"),
                          labels = c("Gain", "Loss")))
frame <- frame[-c(73,113,119,128,173,202,210,280),] 
glm1 <- glm(risky_choice ~ sample*framing, frame, 
            family = binomial)

(a1 <- aov_ez(id = "session_id", dv = "risky_choice", data = frame, 
              between =  c("sample", "framing")))
```

```{r echo=FALSE}
# looking at the data 
afex_plot(a1, "framing", "sample",
          mapping = c("fill", "shape"),
          data_geom = ggpol::geom_boxjitter, 
          data_arg = list(width = 0.5), 
          dodge = 0.7) +
  labs(x = "Frame condition", y = "Probability of risky choice") +
  jtools::theme_apa() +
  theme(legend.position="bottom")
```
  
*Fig 1.*  Probability of risky choice in gain and loss frames by the three samples. The points represent the distribution of data points for each sample. 

Figure 1 suggests that the probability of making a risky choice is lower in the gain frame than the loss frame for Connecticut College and Malaysia, but perhaps not Prague. Testing the model adequacy, we see that the observed data and predictions for the highest-order design cells agree with one another. On the other hand, the predictions for the lower order design cells do not agree with the observed data though they lie in the same direction - there is a lower probability of making risky choices in the gain frame as compared to the loss frame. Considering a non-significant interaction was also found, the significant main effects of framing and sample should be interpreted cautiously, especially sample which has a signifiance close to the boundary at **p = 0.03**.  
  
Nevertheless, the probability of making a risky choice when under the loss frame is significantly higher **(*Pr* = 0.62, *SE* = 0.05)** than when in the gain frame **(*Pr* = 0.37, *SE* = 0.04, p < 0.0001**. The 95% confidence interval of 0.30 to 0.45 for the probability of the gain frame is also below the 0.5 boundary, suggesting that we are 95% confident that a safe choice is more probable in the gain frame. Similarly, the confidence interval of 0.53 to 0.71 for the loss frame suggests we are 95% confident that a risky choice is more probable in the loss grame. Hence, the data does show the usual framing effect of more risky choices made when a loss frame is used than with a grain frame.  
  
#### Do participants generally exhibit risk averse behaviours?
A binomial generalised linear model by was implemented by participant to investigate the probability of risk averse behaviours in participants. An overall probability of choosing the risky option was found at 48.3%. However, the 95% confidence interval for the true overall probability laid between 42.5% and 54.3%. Hence, we cannot be sure that participants are actually risk averse as the probability lies beyond 50%.

#### Do the effects hold across samples?
Participants in Connecticut College had a significantly higher probability of choosing risky choices **(*Pr* = 0.65, *SE* = 0.08)** in the loss condition than when in the gain condition **(*Pr* = 0.255, *SE* = 0.06, p < 0.0001)**. Similarly, participants in Malaysia had a significantly higher probability of choosing risky choices **(*Pr* = 0.744, *SE* = 0.07)** in the loss condition than when in the gain condition **(*Pr* = 0.464, *SE* = 0.07, p < 0.05)**. But there was no significant difference in the probability of choosing risky choices in both gain and loss conditions for participants in Prague **(p > 0.05)**. However, the confidence intervals for these estimated probabilities do not lie below 0.5 for the gain frame in Malaysia and Prague, nor do they lie above 0.5 for the loss frame for Connecticut College and Prague, which would allow us to be confident in the existence of framing effects by sample. Moreover, considering the interaction between sample and framing was non-significant and the lower-order effects do not fully agree with the observed data, the data cannot determine a framing effect exists by sample.  
  
## Section 2  
```{r message = FALSE}
library("tidyverse")
library("afex")
library("emmeans")
theme_set(theme_bw(base_size = 15) + 
            theme(legend.position="bottom", 
                  panel.grid.major.x = element_blank(),
                  panel.grid.minor.x = element_blank()))
options(digits = 3)
set_sum_contrasts()
```

```{r message = FALSE}
# Setting up and preparing data
frame <- read_csv("framing_replication.csv")
str(frame)

# factorising columns
frame <- frame %>% 
  mutate(sample = factor(sample, 
                         levels = c("ConnecticutCollege", 
                                    "Malaysia", 
                                    "Prague")),
         framing = factor(framing,
                          levels = c("gain", "loss"),
                          labels = c("Gain", "Loss")))

contrasts(frame$sample)
contrasts(frame$framing)
which(duplicated(frame))
# returns an array of rows and cols with missing data
which(is.na(frame), arr.ind=T)
# there are 8 cases with missing data in risky_choice that we are removing
frame <- frame[-c(73,113,119,128,173,202,210,280),] 
glm1 <- glm(risky_choice ~ sample*framing, frame, 
            family = binomial)

(a1 <- aov_ez(id = "session_id", dv = "risky_choice", data = frame, 
              between =  c("sample", "framing")))

# looking at the data 
afex_plot(a1, "framing", "sample",
          mapping = c("fill", "shape"),
          data_geom = ggpol::geom_boxjitter, 
          data_arg = list(width = 0.5), 
          dodge = 0.7) +
  labs(x = "Frame condition", y = "Probability of risky choice") +
  jtools::theme_apa() +
  theme(legend.position="bottom")

# checking model adequacy - lower order
frame %>% 
  group_by(framing) %>% 
  summarise(choice = mean(risky_choice))
(em1 <- emmeans(glm1, "framing", type = "response"))

# checking model adequacy - highest order
frame %>% 
  group_by(sample, framing) %>% 
  summarise(choice = mean(risky_choice))
(em2 <- emmeans(glm1, c("sample", "framing"), type = "response"))

glm2 <- glm(risky_choice ~ session_id, frame, 
            family = binomial)
car::Anova(glm2, type = 3)
(em3 <- emmeans(glm2, "session_id", type = "response"))

afex_plot(a1, "sample", "framing",
          mapping = c("fill", "shape"),
          data_geom = ggpol::geom_boxjitter, 
          data_arg = list(width = 0.5), 
          dodge = 0.7) +
  labs(x = "Sample location", y = "Probability of risky choice") +
  jtools::theme_apa() +
  theme(legend.position="bottom")

con1 <- list(
  Congain_Conloss = c(1, 0, 0, -1, 0, 0),
  Malgain_Malloss = c(0, 1, 0, 0, -1, 0),
  Pragain_Praloss = c(0, 0, 1, 0, 0, -1)
)

contrast(em2, con1, adjust = "holm", type = "response")

```

